Le titre

### 2022-09-04 Manuscrit Thèse/chapter 4/Chapter 4.tex
tata

Using NV-NV CR for magnetometry in non-zero magnetic field has already been proposed \citep{akhmedzhanov2017microwave, akhmedzhanov2019magnetometry}. We will focus here on low to zero field magnetometry, which presents practical advantages as the relative orientation of the diamond with the magnetic field does not need to be controlled, but theoretical complications due to the complex physics of the NV centers in low magnetic field.

@@

We will then focus on the low field depolarization magnetometry (LFDM) protocol which exploits the previously studied NV-NV CR at low magnetic field. We will start by giving a short overview of NV ensemble magnetometry before characterizing the LFDM protocol and comparing it to other magnetometry protocols. Finally we will give some perspectives on the how to improve LFDM and on potential applications.

@@

Due to the large zero field splitting $D=2870 MHz$ between the $\ket{0}$ and $\ket{\pm 1}$ states, we will consider the $\ket{0}$ to always be an eigenstate of the spin Hamiltonian under zero external field (which is equivalent to say that we neglect the terms in $\ket{0}\bra{\pm 1}$ in the spin Hamiltonian). The problem is then reduced to the $\{ \ket{-1}, \ket{+1} \}$ subsystem.

@@

Experimentally, almost all our samples show the characteristic two bumps in zero external field ODMR. Given the simulation results, the only parameter that can give rise to this shape is the electric field. We will therefore consider that the NV Hamiltonian of our samples is dominated by the local electric field, and more specifically by the transverse electric field $E_\perp\equiv E_x + i E_y$ given the ratio between $d_\perp$ and $d_\parallel$.

@@

Parler du champs magnétique terrestre.

@@

We will consider here the case of purely transverse magnetic field with respect to the NV axis, i.e. $\mathbf{B}=B_x \hat{e}_x + B_y \hat e_y$, and more specifically the regime where $d_\perp E_\perp < \frac{(\gamma_e B_\perp)^2}{D} << D$. In practice, this generally means $20\ \rm G \lesssim B_\perp \lesssim 200\ \rm G$.

In this regime, the NV Hamiltonian eigenstates are similar to the case dominated by the transverse electric field and can be written $\approx \ket{0}, \ket{\pm}$\citep{qiu2021nuclear, qiu2022nanoscale}, of eigenvalues $\approx -\frac{(\gamma_e B_\perp)^2}{D},D$ and $D+\frac{(\gamma_e B_\perp)^2}{D}$,  where:
\begin{equation}
\ket{\pm}=\frac{\ket{+1}\pm e^{-2i\phi_B}\ket{-1}}{\sqrt{2}},
\end{equation}
and $\tan(\phi_B)=B_y/B_x$.

@@

In conclusion, whenever the spin Hamiltonian is dominated by a transverse field, either electric or magnetic, we can consider that the eigenstates of the spin Hamiltonian or $\ket{0}, \ket{-}$ and $\ket{+}$, whereas when the Hamiltonian is dominated by the longitudinal magnetic field, its eigenstates are $\ket{0}, \ket{-1}$ and $\ket{+1}$.

@@

\begin{table}[htbp]
\centering
\caption{Simulated depolarization rate for flip-flops and double flips in zero magnetic field}
 \label{table double flips}
\begin{tabular}{c|cc}
%\toprule
{} & $l_c=0$ & $l_c=\infty$ \\
\midrule
$\bar{G}^2=1$ & $126.8 \Gamma_0^{\rm th}$ & $126.8 \Gamma_0^{\rm th}$ \\
$\bar{G}^2=0.5$ & $NA \Gamma_0^{\rm th}$ & $NA \Gamma_0^{\rm th}$ \\
$\bar{G}^2=0$ & $51.4 \Gamma_0^{\rm th}$ & $55 \Gamma_0^{\rm th}$ \\
%\bottomrule
\end{tabular}
\end{table}

@@

To quantify its impact in the depolarization rate, we employ eq. [REF], wherche use the values $2\gamma_f = 6.5\ \rm MHz$ and we take for $\Gamma_f=\Gamma_NV$ the values of half width at half maximum found in Fig. \ref{ESR for T2*}.

We then find that the modification of $T_2^*$ should increase the decay rate in zero field by $\sim 6 \%$ for sample Sumi-2, and by $\sim 19 \%$ for sample CVD-pink.

@@

Both AC (up to $\sim$ 10 kHz) and DC-magnetometry are possible with this setup. to improve the sensitivity however, an AC field needs to be added for the DC sensing, and inversely a DC field needs to be added for the AC sensing. Both of these added fields are typically of a few G for optimal performances, which limits the ultra-low field applications.

@@

%Je parle pas de la dépendence en puissance ou du champ transverse, je laisse ça pour plus tard.
\section{Conclusion and perspectives}
We have seen in this chapter that dense ensemble of NV centers spontaneously depolarize at low magnetic field. We have seen that the depolarization is mediated by dipole-dipole coupling between the NV centers.

We have identified three mechanisms by which the depolarization is increased in zero-field: the inter-class flip-flops, the double flips, and the change in the NV eigenstates (which includes the change in $T_2^*$).We have then quantified both theoretically and experimentally the contribution of each of these mechanisms to the zero-field depolarization.

@@

Fig. \ref{applications} illustrates two potential applications for LFDM. The first one is to probe a surface with irregularities in the micron-scale to perform wide-field magnetic imaging. The current leading technology for wide field imaging is to use a bulk CVD diamond with a very well polished surface in direct contact with the probed sample \citep{levine2019principles, scholten2021widefield}. While the diamond surface can be controlled within a few nm, the same is not always true for the sample, especially for paleomagnetism or biomagnetism. And even for two perfectly flat surface, controlling the angle between the diamond and the sample can be challenging.

Instead we propose to deposit a thin layer of 1 $\mu$m diamonds on top of the sample in order to always be within $\sim$ 1 $\mu$m of the sample. Going below 1 $\mu$m may not prove useful as the spatial resolution is ultimately limited by the diffraction limit of the lens. One advantage of this method is that commercially available samples (Adamas Nano 1 $\mu$ m red fluorescent diamond) are well suited for this task. I have measured the sensitivity on some of these samples and found the same volumetric sensitivity $\eta_v \sim 6 \ \rm{\mu T}\mu \rm{m}^{3/2} \rm{Hz}^{-1/2}$ than for the 15 $\mu$m samples. This sensitivity is already better than some previous applications of wide field imaging at $20  \ \rm{\mu T}\mu \rm{m} \rm{Hz}^{-1/2}$ \citep{glenn2017micrometer}. Another potential advantage of LFDM is that a homogeneous magnetic field or microwave intensity along the sample is not required. Only a $\sim 10 \ \rm G$ AC magnetic field is required, and the sensitivity does not depend critically on the AC field homogeneity.

>>>>

Using NV-NV CR for magnetometry in non-zero magnetic field has already been proposed \citep{akhmedzhanov2017microwave, akhmedzhanov2019magnetometry}. We will focus here on the low to zero magnetic field region, which presents practical advantages as the relative orientation of the diamond with the magnetic field does not need to be controlled, but theoretical complications due to the more complex physics of the NV centers under low magnetic field.

@@

We will then focus on the low field depolarization magnetometry (LFDM) protocol which exploits the previously studied NV-NV CR at low magnetic field. We will start by giving a short overview of NV ensemble magnetometry before characterizing the LFDM protocol and comparing it to other magnetometry protocols. Finally we will give some perspectives on how to improve LFDM and on potential applications.

@@

Due to the large zero field splitting $D=2870\ \rm MHz$ between the $\ket{0}$ and $\ket{\pm 1}$ states, we will consider the $\ket{0}$ to always be an eigenstate of the spin Hamiltonian under zero external field (which is equivalent to say that we neglect the terms in $\ket{0}\bra{\pm 1}$ in the spin Hamiltonian). The problem is then reduced to the $\{ \ket{-1}, \ket{+1} \}$ subsystem.

@@

Experimentally, almost all our samples show the characteristic two bumps in zero external field ODMR. Given the simulation results, the only parameter that can give rise to this shape is the electric field.

We should note that another potential explanation for the splitting of the $\ket{\pm 1}$ states could be the presence of a residual magnetic field, not randomly distributed among all the spins but homogeneous on the whole sample. In particular, we do not shield for the earth magnetic field in our experiment. The splitting caused by the earth magnetic field has a value $\gamma B_{\rm earth} \approx 1\ \rm MHz$ is $\sim 4$ times weaker than the splitting we typically observe.

We will therefore consider that the low external magnetic field spin Hamiltonian of our samples is dominated by the local electric field, and more specifically by the transverse electric field $E_\perp\equiv E_x + i E_y$ given the ratio between $d_\perp$ and $d_\parallel$.

@@

@@

We now turn to the case of purely transverse magnetic field with respect to the NV axis, i.e. $\mathbf{B}=B_x \hat{e}_x + B_y \hat e_y$, and more specifically to the regime where $d_\perp E_\perp < \frac{(\gamma_e B_\perp)^2}{D} << D$. In practice, this generally corresponds to $20\ \rm G \lesssim B_\perp \lesssim 200\ \rm G$.

In this regime, the NV Hamiltonian eigenstates are similar to the case where it is dominated by the transverse electric field and can be written $\approx \ket{0}, \ket{\pm}$\citep{qiu2021nuclear, qiu2022nanoscale}, of eigenvalues $\approx -\frac{(\gamma_e B_\perp)^2}{D},D$ and $D+\frac{(\gamma_e B_\perp)^2}{D}$,  where:
\begin{equation}
\ket{\pm}=\frac{\ket{+1}\pm e^{-2i\phi_B}\ket{-1}}{\sqrt{2}},
\end{equation}
and $\tan(\phi_B)=B_y/B_x$.

@@

In conclusion, whenever the spin Hamiltonian is dominated by a transverse field, either electric or magnetic, we can consider that the eigenstates of the spin Hamiltonian to be $\ket{0}, \ket{-}$ and $\ket{+}$, whereas when the Hamiltonian is dominated by the longitudinal magnetic field, its eigenstates are $\ket{0}, \ket{-1}$ and $\ket{+1}$.

@@

\begin{table}[htbp]
\centering
\caption{Simulated depolarization rate for flip-flops and double flips in zero magnetic field}
 \label{table double flips}
\begin{tabular}{c|cc}
%\toprule
{} & $l_c=0$ & $l_c=\infty$ \\
\midrule
$\bar{G}^2=1$ & $126.8 \Gamma_0^{\rm th}$ & $126.8 \Gamma_0^{\rm th}$ \\
$\bar{G}^2=0.5$ & $93.7 \Gamma_0^{\rm th}$ & $95.4 \Gamma_0^{\rm th}$ \\
$\bar{G}^2=0$ & $51.4 \Gamma_0^{\rm th}$ & $55 \Gamma_0^{\rm th}$ \\
%\bottomrule
\end{tabular}
\end{table}

@@

To quantify its impact in the depolarization rate, we employ eq. [REF], where we use the values $2\gamma_f = 6.5\ \rm MHz$ estimated in the last chapter and we take for $\Gamma_f=\Gamma_{\rm NV}$ the values of half width at half maximum found in Fig. \ref{ESR for T2*}.

We then find that the modification of $T_2^*$ should increase the decay rate in zero field by $\sim 9 \%$ for sample Sumi-2, and by $\sim 24 \%$ for sample CVD-pink.

@@

Both AC (up to $\sim$ 10 kHz) and DC-magnetometry are possible with this setup. To improve the sensitivity however, an AC field needs to be added for the DC sensing, and inversely a DC field needs to be added for the AC sensing. Both of these added fields are typically of a few G for optimal performances, which limits the ultra-low field applications.

@@

%Je parle pas de la dépendence en puissance ou du champ transverse, je laisse ça pour plus tard.
\section{Conclusion and perspectives}
We have seen in this chapter that dense ensemble of NV centers spontaneously depolarize at low magnetic field, and wee have seen that the depolarization is mediated by dipole-dipole coupling between the NV centers. We have identified three mechanisms by which the depolarization is increased in zero-field: the inter-class flip-flops, the double flips, and the change in the NV eigenstates (which includes the change in $T_2^*$).We have then quantified both theoretically and experimentally the contribution of each of these mechanisms to the zero-field depolarization.

@@

Fig. \ref{applications} illustrates two potential applications for LFDM. The first one is to probe a surface with irregularities in the micron-scale to perform wide-field magnetic imaging. The current leading technology for wide field imaging is to use a bulk CVD diamond with a very well polished surface in direct contact with the probed sample \citep{levine2019principles, scholten2021widefield}. While the diamond surface can be controlled within a few nm, the same is not always true for the sample, especially for paleomagnetism or biomagnetism samples. Furthermore, controlling the angle between the diamond and the sample can be challenging even for two perfectly flat surface.

Instead we propose to deposit a thin layer of 1 $\mu$m diamonds on top of the sample in order to always always have a sensor within $\sim$ 1 $\mu$m of the sample. Going below 1 $\mu$m may not prove useful as the spatial resolution is ultimately limited by the diffraction limit of the lens. One advantage of this method is that commercially available samples (Adamas Nano 1 $\mu$ m red fluorescent diamond) are well suited for this task. I have measured the sensitivity on some of these samples and found the same volumetric sensitivity $\eta_v \sim 6 \ \rm{\mu T}\mu \rm{m}^{3/2} \rm{Hz}^{-1/2}$ than for the 15 $\mu$m samples. This sensitivity is already better than some previous applications of wide field imaging at $20  \ \rm{\mu T}\mu \rm{m} \rm{Hz}^{-1/2}$ \citep{glenn2017micrometer}. Another potential advantage of LFDM is that a homogeneous magnetic field or microwave intensity along the sample is not required. Only a $\sim 10 \ \rm G$ AC magnetic field is required, and the sensitivity does not depend critically on the AC field homogeneity.
